---
title: "Cbf_analysis_stats"
author: "Kyle Drover"
date: "2023-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up

Loading libraries and setting up functions

```{r}
library(remotes)
install_github("cran/estimability")
```


```{r}
if(!require(Hmisc))install.packages("Hmisc")
if(!require(multcomp))install.packages("multcomp")
if(!require(emmeans))install.packages("emmeans")
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(Hmisc)
library(multcomp)
library(emmeans)


genotype_names <- list(
  "WT" = expression(bolditalic("Zic3")^bolditalic("+/+")),
  "Mut" = expression(bolditalic("Zic3")^bolditalic("Ka/Y"))
)

# looks complicated, cause it is. It a function that labels things well 
genotype_labeller <- function(variable,value){return(genotype_names[value])}
```

## reading and arranging data

You can also embed plots, for example:

```{r}
Katuns <- c("Ka96_e3", "Ka109_e4", "Ka109_e5")
cbf_data <- read_csv("Cilarrmove_results_for_plotting.csv") %>% mutate(Genotype = ifelse((Embryo_Id %in% Katuns),"Mut", "WT"))

cbf_data$Video_No. <- as.character(cbf_data$Video_No.)

cbf_data <- arrange(transform(cbf_data, Genotype=factor(Genotype,levels=c("WT","Mut"))))
```

## Part 1 plotting histograms:

Before removing 0-1 (i.e., spaces where the values were negative space which is highly dependent on what orientation the cilia is at)
```{r, fig.height=15}
ggplot(cbf_data, aes(x=Frequency..Hz., y=Percentage...., col=interaction(Video_No., Embryo_Id)))+
  geom_line()+
  facet_wrap(~Embryo_Id, ncol=1)



ggplot(cbf_data, aes(x=Frequency..Hz., y=Percentage...., col=Embryo_Id))+
  geom_line()+
  facet_wrap(~Embryo_Id, ncol=1)

ggplot(cbf_data, aes(x=Frequency..Hz., y=Percentage...., col=Genotype))+
  geom_bar(stat="identity")+
  facet_wrap(~Genotype, ncol=1)
```

```{r}
# define colour scheme
WT_shades <- "#425AA7"
Mut_shades <- "#BA89AC"

embryo_shades <- c("#425AA7","#425AA7","#BA89AC","#BA89AC","#425AA7","#425AA7","#425AA7","#BA89AC")
embryo_shades_WM <- c("#425AA7","#425AA7","#425AA7","#425AA7","#425AA7","#BA89AC","#BA89AC","#BA89AC")
```


After removing 0-1:
```{r, fig.height=7, fig.width=7}

cbf_cleaned <- cbf_data %>% filter(Frequency..Hz. > 1)

# individual cilia frequency plots
Indiv_hz <-   ggplot(cbf_cleaned, aes(x=Frequency..Hz., y=Percentage...., col=interaction(Video_No., Embryo_Id)))+
              geom_line()+
              facet_wrap(~Embryo_Id, ncol=1)

# average frequency plot per embryo
Embryo_hz <-  ggplot(cbf_cleaned, aes(x=Frequency..Hz., y=Percentage...., col=Embryo_Id))+
              geom_line()+
              facet_wrap(~Embryo_Id, ncol=1)+
              scale_color_manual(values = embryo_shades) 

# average frequency per genotype
Genotype_hz <-  ggplot(cbf_cleaned, aes(x=Frequency..Hz., y=Percentage...., col=Genotype))+
                geom_bar(stat="identity")+
                facet_wrap(~Genotype, ncol=1)+
                scale_color_manual(values = c(WT_shades, Mut_shades)) 


Indiv_hz
Embryo_hz
Genotype_hz

```

```{r}
ggsave("Embryo_cilia_hz.png", plot = Embryo_hz, dpi=300, height = 7, width = 7, units = "in")
ggsave("Ind_cilia_hz.png", plot = Indiv_hz, dpi=300, height = 7, width = 7, units = "in")
ggsave("Genotype_cilia_hz.png", plot = Genotype_hz, dpi=300, height = 7, width = 7, units = "in")

```


## Part 2- Plotting weighted mean + S.E.M


``` {r}
# groups data per variable and then calculates the wieghted average + S.E.M
weighted_mean_per_embryo <- cbf_cleaned %>%
                            group_by(Embryo_Id) %>%
                            dplyr::summarize(weighted_mean_value = weighted.mean(Frequency..Hz., w = Percentage....),
                                      weighted_var_value = wtd.var(Frequency..Hz.,Percentage....),
                                      weighted_n = sum(Percentage....),
                                      weighted_sem_value = sqrt(weighted_var_value / weighted_n))


weighted_mean_per_embryo_per_vid <- cbf_cleaned %>%
                                    group_by(Embryo_Id, Video_No.) %>%
                                    dplyr::summarize(weighted_mean_value = weighted.mean(Frequency..Hz., w = Percentage....),
                                                     weighted_var_value = wtd.var(Frequency..Hz.,Percentage....),
                                                     weights = sum( (Percentage.... / sum(Percentage....))^2 ),
                                                     weighted_sem_value = sqrt(weighted_var_value* weights))


weighted_mean_per_genotype <- cbf_cleaned %>%
                              group_by(Genotype) %>%
                              dplyr::summarize(weighted_mean_value = weighted.mean(Frequency..Hz., w = Percentage....),
                                               weighted_var_value = wtd.var(Frequency..Hz.,Percentage....),
                                               weights = sum( (Percentage.... / sum(Percentage....))^2 ),
                                               weighted_sem_value = sqrt(weighted_var_value* weights))
```
  
```{r, fig.height=2.5, fig.width=4}  

# plot individual videos
WMbar_ind <-  ggplot(weighted_mean_per_embryo_per_vid, aes(x = interaction(Video_No., Embryo_Id), y = weighted_mean_value, fill=Embryo_Id))               +
              geom_bar(stat = "identity") +
              geom_errorbar(aes(ymin = weighted_mean_value - weighted_sem_value, ymax = weighted_mean_value + weighted_sem_value),
                            width = 0.2, color = "black") +
              scale_x_discrete(guide = guide_axis(n.dodge=6))+
              labs(title = "Weighted Mean and WSE per Embryo",
                   x = "Each Individual Video",
                   y = "Weighted Mean") +
              theme_minimal()

# plot average per embryo
WMbar_embryo <- ggplot(weighted_mean_per_embryo, aes(x = Embryo_Id, y = weighted_mean_value, fill=Embryo_Id)) +
                geom_bar(stat = "identity") +
                geom_errorbar(aes(ymin = weighted_mean_value - weighted_sem_value, ymax = weighted_mean_value + weighted_sem_value),
                              width = 0.2, color = "black") +
                labs(title = "Weighted Mean and WSE per Embryo",
                     x = "Embryo Id",
                     y = "Weighted Mean") +
                theme_minimal()+
                scale_fill_manual(labels = unlist(genotype_names), values=embryo_shades_WM)


# plot average per genotype
WMbar_genotype <- ggplot(weighted_mean_per_genotype, aes(x = Genotype, y = weighted_mean_value, fill=Genotype)) +
                geom_bar(stat = "identity") +
                geom_errorbar(aes(ymin = weighted_mean_value - weighted_sem_value, ymax = weighted_mean_value + weighted_sem_value),
                              width = 0.2, color = "black") +
                labs(title = "Weighted Mean and WSE per Genotype",
                     x = "Genotype",
                     y = "Weighted Mean") +
                scale_fill_manual(labels = unlist(genotype_names), values=c(WT_shades, Mut_shades))+
                scale_x_discrete(labels = unlist(genotype_names))+
                theme_minimal()

WMbar_ind
WMbar_embryo
WMbar_genotype
```


```{r}
ggsave("Embryo_cilia_WM.png", plot = WMbar_embryo, dpi=300, height = 2.5, width = 4, units = "in")
ggsave("Ind_cilia_WM.png", plot = WMbar_ind, dpi=300, height = 2.5, width = 4, units = "in")
ggsave("Genotype_cilia_WM.png", plot = WMbar_genotype, dpi=300, height = 2.5, width = 4, units = "in")
```

## Part 3 Stats:

Just comparing Genotypes:

```{r}

data_for_stats <- cbf_cleaned

data_for_stats$Embryo_Id <- as.factor(data_for_stats$Embryo_Id)


genotype_test <- aov(Frequency..Hz.~Genotype, data=data_for_stats)


summary(genotype_test)

```


Comparing individual embryos
```{r}

# perform a ANOVA across all embryos to show a statistical difference and then go and do the post-hoc test
Embryo_Id_test <- aov(Frequency..Hz.~Embryo_Id, data=data_for_stats)
summary(Embryo_Id_test)


```


```{r}
#So there's a couple of different ways of doing the contrasts. One is using Tukey's
posthoc <- glht(Embryo_Id_test, linfct=mcp(Embryo_Id="Tukey"))
cat("\n")
cld(posthoc)
```
```{r}
#the others is using eemenas
ID_test <- emmeans(Embryo_Id_test, "Embryo_Id")

cld(ID_test)

```



